---
title: "Chapter 3"
author: "Bob"
date: "2025-05-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries, message=FALSE, warning=FALSE}
# libraries
library('rjags')      # coda.samples
library('jagsUI')     # wrapper for rjags
library('mcmcplots')  # caterplot
library('coda')
library('knitr')
library('dplyr')
library('tidyr')
library('ggplot2')
library('knitr')      # HTML table display

source('NMA_utils.R')
```

## Notes

# `Ch3_thromboFE_criticism`

full Thrombolitics data: 50 trials, 9 treatments
==============================  
1	SK
2	t-PA
3	Acc t-PA
4	SK + t-PA
5	r-PA
6	TNK
7	PTCA
8	UK
9	ASPAC
===============================

* Binomial likelihood, logit link
* Fixed effect model

This code is part of Dias, Ades, Welton, Jansen and Sutton (2018) Network Meta-Analysis for Decision Making. 
This work should be cited whenever the code is used whether in its standard form or adapted.

```{r Ch3_thromboFE_criticism}
model_code <- "model{						# *** PROGRAM STARTS
for(i in 1:ns){					# LOOP THROUGH STUDIES
     mu[i] ~ dnorm(0,.0001)			# vague priors for all trial baselines
     for (k in 1:na[i])  {				#  LOOP THROUGH ARMS
          r[i,k] ~ dbin(p[i,k],n[i,k])			# binomial likelihood
          logit(p[i,k]) <- mu[i] + d[t[i,k]] - d[t[i,1]]	# model for linear predictor
          rhat[i,k] <- p[i,k] * n[i,k]				# expected value of the numerators 
          dev[i,k] <- 2 * (r[i,k] * (log(r[i,k])-log(rhat[i,k]))     	#Deviance contribution
             +  (n[i,k]-r[i,k]) * (log(n[i,k]-r[i,k]) - log(n[i,k]-rhat[i,k])))   
        }
     resdev[i] <- sum(dev[i,1:na[i]])			# summed residual deviance contribution for this trial
     }   
totresdev <- sum(resdev[])				#Total Residual Deviance 
d[1]<-0					# treatment effect is zero for reference treatment
for (k in 2:nt){  d[k] ~ dnorm(0,.0001) }		# vague priors for treatment effects

}						# *** PROGRAM ENDS
" %>% textConnection


# !!! I made study_ID a regular column instead of comments
data_df <- "r[,1]	n[,1]	r[,2]	n[,2]	r[,3]	n[,3]	t[,1]	t[,2]	t[,3]	na[]	study_ID
1472	20251	652	10396	723	10374	1	3	4	3	1
9	130	6	123	NA	NA	1	2	NA	2	2
5	63	2	59	NA	NA	1	2	NA	2	3
3	65	3	64	NA	NA	1	2	NA	2	4
887	10396	929	10372	NA	NA	1	2	NA	2	5
1455	13780	1418	13746	1448	13773	1	2	9	3	6
7	85	4	86	NA	NA	1	2	NA	2	7
12	159	7	157	NA	NA	1	2	NA	2	8
10	135	5	135	NA	NA	1	2	NA	2	9
4	107	6	109	NA	NA	1	4	NA	2	10
285	3004	270	3006	NA	NA	1	5	NA	2	11
11	149	2	152	NA	NA	1	7	NA	2	12
1	50	3	50	NA	NA	1	7	NA	2	13
8	58	5	54	NA	NA	1	7	NA	2	14
1	53	1	47	NA	NA	1	7	NA	2	15
4	45	0	42	NA	NA	1	7	NA	2	16
14	99	7	101	NA	NA	1	7	NA	2	17
9	41	3	46	NA	NA	1	7	NA	2	18
42	421	29	429	NA	NA	1	7	NA	2	19
2	44	3	46	NA	NA	2	7	NA	2	20
13	200	5	195	NA	NA	2	7	NA	2	21
2	56	2	47	NA	NA	2	7	NA	2	22
3	55	1	55	NA	NA	3	7	NA	2	23
10	94	3	95	NA	NA	3	7	NA	2	24
40	573	32	565	NA	NA	3	7	NA	2	25
2	61	3	62	NA	NA	3	7	NA	2	26
16	419	20	421	NA	NA	3	7	NA	2	27
5	69	3	71	NA	NA	3	7	NA	2	28
5	75	5	75	NA	NA	3	7	NA	2	29
59	782	52	790	NA	NA	3	7	NA	2	30
5	81	2	81	NA	NA	3	7	NA	2	31
16	226	12	225	NA	NA	3	7	NA	2	32
8	66	6	71	NA	NA	3	7	NA	2	33
522	8488	523	8461	NA	NA	3	6	NA	2	34
356	4921	757	10138	NA	NA	3	5	NA	2	35
13	155	7	169	NA	NA	3	5	NA	2	36
10	203	7	198	NA	NA	1	8	NA	2	37
3	58	2	52	NA	NA	1	9	NA	2	38
3	86	6	89	NA	NA	1	9	NA	2	39
3	58	2	58	NA	NA	1	9	NA	2	40
13	182	11	188	NA	NA	1	9	NA	2	41
2	26	7	54	NA	NA	3	8	NA	2	42
12	268	16	350	NA	NA	3	8	NA	2	43
5	210	17	211	NA	NA	3	9	NA	2	44
3	138	13	147	NA	NA	3	9	NA	2	45
8	132	4	66	NA	NA	2	8	NA	2	46
10	164	6	166	NA	NA	2	8	NA	2	47
6	124	5	121	NA	NA	2	8	NA	2	48
13	164	10	161	NA	NA	2	9	NA	2	49
7	93	5	90	NA	NA	2	9	NA	2	50
" %>% 
  load_dataframe_from_string

my_data <- data_df %>%
  load_matrix_data_from_df %>%
  append(list(
    ns = nrow(data_df), # number of studies
    nt = 9  # number of treatments
  ))


initial_values <- list(
  # chain 1
  list(
  d=c(NA,0,0,0,0,	0,0,0,0),
  mu=c(0,0,0,0,0,  0,0,0,0,0,  0,0,0,0,0,  0,0,0,0,0,  0,0,0,0,0,
  0,0,0,0,0,  0,0,0,0,0,  0,0,0,0,0,  0,0,0,0,0,  0,0,0,0,0)
  ),
  
  # chain 2
  list(
  d=c(NA,.1,-1,-.2, -.2,	-.2,-.2,-.2,-.2),
  mu=c(1,-1,-2,0,0,    -2,1,0,2,2,    1,-1,-2,0,0,   -2,1,0,2,2,   -2,-0.5,-3,0.5,-1,
  1,-1,-2,0,0,    -2,1,0,2,2,    1,-1,-2,0,0,   -2,1,0,2,2,   -2,-0.5,-3,0.5,-1)
  ),
  
  # chain 3
  list(d=c( NA, 2,5,-3,1,     -7,4,2,-1), mu=c(-3, 5,-1,-3,7,      -3, -4, -3, -3, 9,    -3, -3,-4, 3, 5,      -3, -2, 1, -3, -7,      -3,5,-1,-3,7,     -3,-4,-3,-3,0,     -3, 5,-1,-3,7,     -3, 5,-1,-3,7,      -3, -4, -3, -3, 9,    -3, -3,-4, 3,1))
)

jags_results <- jags(
  data = my_data,
  inits = initial_values,
  parameters.to.save = c("d", "dev"),
  model.file = model_code,
  n.chains = length(initial_values),
  n.adapt = 100,
  n.iter = 60000,
  n.burnin = 10000
)

jags_summary <- summary(jags_results)


jags_summary[,c('mean', 'sd', '2.5%', '50%', '97.5%')] %>% 
  knitr::kable(., format='html', escape=FALSE, caption="JAGS Results")

```

```{r Ch3_thromboFE_criticism_compare_to_WinBUGS, echo=FALSE}

winbugs_summary <- "node	 mean	 sd	 MC error	2.5%	median	97.5%	start	sample
d[2]	0.002561	0.03029	2.116E-4	-0.05732	0.002636	0.06122	40001	80000
d[3]	-0.1772	0.04343	3.68E-4	-0.2627	-0.1771	-0.09245	40001	80000
d[4]	-0.04959	0.04666	2.371E-4	-0.1416	-0.04948	0.04154	40001	80000
d[5]	-0.1237	0.06025	5.175E-4	-0.2424	-0.1236	-0.006041	40001	80000
d[6]	-0.172	0.07694	6.04E-4	-0.3235	-0.1719	-0.02118	40001	80000
d[7]	-0.4761	0.1009	5.911E-4	-0.676	-0.4759	-0.2781	40001	80000
d[8]	-0.2033	0.2214	0.00148	-0.6371	-0.2031	0.2294	40001	80000
d[9]	0.01619	0.03678	2.303E-4	-0.05624	0.01626	0.08838	40001	80000
dev[1,1]	0.9858	1.381	0.00755	9.463E-4	0.4549	4.908	49201	61600
dev[1,2]	0.9552	1.334	0.008188	9.862E-4	0.4369	4.771	49201	61600
dev[1,3]	1.001	1.414	0.005665	9.303E-4	0.4561	4.964	49201	61600
dev[2,1]	0.818	1.091	0.004457	0.00102	0.4002	3.89	49201	61600
dev[2,2]	0.6716	0.8998	0.003557	6.988E-4	0.3284	3.22	49201	61600
dev[3,1]	1.226	1.495	0.006319	0.001611	0.685	5.368	49201	61600
dev[3,2]	1.017	1.177	0.004897	0.001717	0.6099	4.203	49201	61600
dev[4,1]	0.5147	0.7259	0.002972	4.733E-4	0.2359	2.57	49201	61600
dev[4,2]	0.5143	0.7254	0.002974	4.724E-4	0.2365	2.589	49201	61600
dev[5,1]	1.212	1.552	0.008877	0.001409	0.6288	5.543	49201	61600
dev[5,2]	1.225	1.548	0.00653	0.001439	0.6435	5.543	49201	61600
dev[6,1]	1.156	1.547	0.01123	0.001212	0.5572	5.567	49201	61600
dev[6,2]	0.9736	1.354	0.005258	8.887E-4	0.4542	4.801	49201	61600
dev[6,3]	0.9965	1.402	0.005835	9.991E-4	0.452	4.998	49201	61600
dev[7,1]	1.073	1.327	0.005368	0.001368	0.5894	4.755	49201	61600
dev[7,2]	0.8836	1.114	0.004233	0.001136	0.4659	3.968	49201	61600
dev[8,1]	1.253	1.474	0.006023	0.001963	0.7321	5.267	49201	61600
dev[8,2]	1.127	1.29	0.005172	0.002054	0.6894	4.635	49201	61600
dev[9,1]	1.469	1.631	0.006335	0.002893	0.9367	5.886	49201	61600
dev[9,2]	1.353	1.429	0.00549	0.003547	0.9051	5.17	49201	61600
dev[10,1]	0.669	0.9073	0.003524	6.945E-4	0.3173	3.227	49201	61600
dev[10,2]	0.8323	1.103	0.004643	9.313E-4	0.415	3.94	49201	61600
dev[11,1]	0.9678	1.348	0.00776	9.104E-4	0.4472	4.807	49201	61600
dev[11,2]	1.002	1.358	0.006556	0.00107	0.4738	4.841	49201	61600
dev[12,1]	2.005	2.149	0.008287	0.004546	1.318	7.732	49201	61600
dev[12,2]	2.72	1.84	0.007188	0.2035	2.379	7.201	49201	61600
dev[13,1]	1.418	1.53	0.005939	0.003455	0.9301	5.482	49201	61600
dev[13,2]	1.874	1.837	0.007689	0.007775	1.36	6.691	49201	61600
dev[14,1]	0.622	0.8751	0.003567	6.341E-4	0.285	3.101	49201	61600
dev[14,2]	0.4191	0.59	0.002405	4.278E-4	0.1925	2.091	49201	61600
dev[15,1]	0.6113	0.8678	0.003578	5.874E-4	0.2758	3.055	49201	61600
dev[15,2]	0.6434	0.8653	0.003547	6.683E-4	0.3135	3.099	49201	61600
dev[16,1]	1.812	2.107	0.008527	0.002832	1.098	7.515	49201	61600
dev[16,2]	3.062	1.554	0.006304	0.8308	2.797	6.799	49201	61600
dev[17,1]	0.8429	1.156	0.005009	8.683E-4	0.3974	4.087	49201	61600
dev[17,2]	0.6412	0.8601	0.003431	6.758E-4	0.3105	3.083	49201	61600
dev[18,1]	1.457	1.691	0.00657	0.002353	0.8725	6.024	49201	61600
dev[18,2]	1.448	1.429	0.00536	0.005375	1.038	5.2	49201	61600
dev[19,1]	0.662	0.9367	0.003983	6.513E-4	0.3005	3.367	49201	61600
dev[19,2]	0.5389	0.7565	0.003358	5.391E-4	0.2486	2.678	49201	61600
dev[20,1]	0.8233	1.1	0.004193	8.638E-4	0.4044	3.945	49201	61600
dev[20,2]	1.087	1.284	0.005508	0.001674	0.6344	4.616	49201	61600
dev[21,1]	1.061	1.396	0.005799	0.001117	0.5335	5.045	49201	61600
dev[21,2]	0.8983	1.053	0.004266	0.001399	0.5295	3.761	49201	61600
dev[22,1]	0.6861	0.9716	0.003997	6.147E-4	0.3064	3.459	49201	61600
dev[22,2]	0.7717	0.9673	0.003746	9.489E-4	0.4145	3.489	49201	61600
dev[23,1]	0.99	1.314	0.00514	0.001129	0.4857	4.685	49201	61600
dev[23,2]	0.6231	0.8248	0.003242	6.861E-4	0.3098	2.936	49201	61600
dev[24,1]	1.718	1.892	0.007411	0.003359	1.099	6.773	49201	61600
dev[24,2]	1.816	1.596	0.006393	0.01564	1.423	5.858	49201	61600
dev[25,1]	0.6579	0.9375	0.003737	6.524E-4	0.2974	3.308	49201	61600
dev[25,2]	0.5967	0.8335	0.003432	6.075E-4	0.2752	2.963	49201	61600
dev[26,1]	0.7015	0.9663	0.003776	6.928E-4	0.3266	3.458	49201	61600
dev[26,2]	0.9233	1.163	0.004651	9.694E-4	0.4903	4.14	49201	61600
dev[27,1]	1.606	1.736	0.007735	0.003611	1.048	6.256	49201	61600
dev[27,2]	1.868	1.825	0.00718	0.005619	1.344	6.576	49201	61600
dev[28,1]	0.6976	0.9744	0.003978	6.649E-4	0.3222	3.501	49201	61600
dev[28,2]	0.4799	0.6788	0.002692	4.16E-4	0.2194	2.414	49201	61600
dev[29,1]	0.6125	0.8758	0.003498	5.996E-4	0.2764	3.111	49201	61600
dev[29,2]	0.6335	0.8659	0.003327	7.156E-4	0.3009	3.094	49201	61600
dev[30,1]	0.9006	1.245	0.005333	8.805E-4	0.4247	4.43	49201	61600
dev[30,2]	0.9332	1.212	0.004605	0.001065	0.4703	4.346	49201	61600
dev[31,1]	0.9772	1.294	0.005354	0.001152	0.4837	4.609	49201	61600
dev[31,2]	0.6999	0.9017	0.003614	7.857E-4	0.3584	3.215	49201	61600
dev[32,1]	0.6	0.8503	0.003451	5.819E-4	0.2713	3.05	49201	61600
dev[32,2]	0.4725	0.6692	0.002766	4.586E-4	0.2154	2.377	49201	61600
dev[33,1]	0.5987	0.843	0.003278	5.549E-4	0.2732	2.975	49201	61600
dev[33,2]	0.4789	0.6812	0.002644	4.436E-4	0.2172	2.41	49201	61600
dev[34,1]	0.9917	1.411	0.00775	9.332E-4	0.4486	4.964	49201	61600
dev[34,2]	0.9943	1.417	0.006099	9.13E-4	0.4491	5.04	49201	61600
dev[35,1]	0.8397	1.174	0.007355	8.112E-4	0.388	4.128	49201	61600
dev[35,2]	0.9179	1.29	0.005019	9.053E-4	0.4244	4.557	49201	61600
dev[36,1]	2.043	1.942	0.007608	0.009034	1.512	7.061	49201	61600
dev[36,2]	1.896	1.788	0.006999	0.009296	1.412	6.5	49201	61600
dev[37,1]	0.721	1.0	0.004376	7.397E-4	0.3349	3.589	49201	61600
dev[37,2]	0.5634	0.8078	0.003445	5.952E-4	0.2515	2.854	49201	61600
dev[38,1]	0.6833	0.9614	0.004037	6.767E-4	0.3114	3.382	49201	61600
dev[38,2]	0.488	0.6883	0.002796	4.648E-4	0.2219	2.448	49201	61600
dev[39,1]	0.8657	1.069	0.004224	0.001196	0.4704	3.797	49201	61600
dev[39,2]	1.074	1.343	0.005449	0.001345	0.5748	4.843	49201	61600
dev[40,1]	0.7234	0.996	0.00412	7.423E-4	0.3379	3.541	49201	61600
dev[40,2]	0.5396	0.7602	0.003158	5.353E-4	0.2461	2.721	49201	61600
dev[41,1]	0.6912	0.9482	0.004047	6.794E-4	0.3259	3.362	49201	61600
dev[41,2]	0.6195	0.854	0.003353	6.43E-4	0.2906	3.033	49201	61600
dev[42,1]	0.6862	0.871	0.004628	8.551E-4	0.3608	3.1	49201	61600
dev[42,2]	0.9792	1.343	0.005448	0.001059	0.4612	4.818	49201	61600
dev[43,1]	0.6154	0.8767	0.003946	5.667E-4	0.2788	3.132	49201	61600
dev[43,2]	0.7288	1.021	0.00424	7.015E-4	0.3349	3.622	49201	61600
dev[44,1]	3.397	2.216	0.009499	0.2857	3.014	8.763	49201	61600
dev[44,2]	2.724	2.459	0.009951	0.0184	2.101	9.061	49201	61600
dev[45,1]	3.284	2.129	0.008799	0.3079	2.903	8.398	49201	61600
dev[45,2]	2.511	2.386	0.009952	0.0113	1.865	8.7	49201	61600
dev[46,1]	0.7436	1.052	0.004298	7.178E-4	0.3345	3.753	49201	61600
dev[46,2]	0.4942	0.6673	0.002834	5.001E-4	0.2345	2.392	49201	61600
dev[47,1]	0.9085	1.218	0.005542	0.001015	0.4386	4.328	49201	61600
dev[47,2]	0.7139	0.9959	0.004013	7.156E-4	0.327	3.551	49201	61600
dev[48,1]	0.609	0.8522	0.003803	6.412E-4	0.2785	3.049	49201	61600
dev[48,2]	0.5345	0.7458	0.003003	4.995E-4	0.2467	2.673	49201	61600
dev[49,1]	0.7606	1.023	0.004048	7.442E-4	0.3654	3.65	49201	61600
dev[49,2]	0.6582	0.9052	0.003522	6.367E-4	0.3106	3.237	49201	61600
dev[50,1]	0.7373	1.009	0.004202	7.702E-4	0.3458	3.595	49201	61600
dev[50,2]	0.5951	0.8291	0.003306	6.19E-4	0.2754	2.937	49201	61600
totresdev	105.9	10.9	0.04228	86.43	105.3	129.2	40001	80000" %>% 
  load_dataframe_from_string

# ???
# bar = post.mean of -2logL; Dhat = -2LogL at post.mean of stochastic nodes
# 	Dbar	Dhat	pD	DIC	
# r	535.557	477.864	57.693	593.250	
# total	535.557	477.864	57.693	593.250	

winbugs_summary %>%
  knitr::kable(., format='html', escape=FALSE, caption="WinBUGS Results")

```

## To Do:

* Tidy up code indentation.
* Sort the JAGS results summary to match the WinBugs summary.
* What are `Dbar` and `r` (see commented out lines near end of preceding chunk)? Are we getting the correct values?



# `Ch3_thromboRE_criticism`

Full Thrombolytics example

* Binomial likelihood, logit link
* Random effects model for multi-arm trials

This code is part of Dias, Ades, Welton, Jansen and Sutton (2018) Network Meta-Analysis for Decision Making. 
This work should be cited whenever the code is used whether in its standard form or adapted.


```{r Ch3_thromboRE_criticism}
model_code <- "model{					# *** PROGRAM STARTS
for(i in 1:ns){				# LOOP THROUGH STUDIES
     w[i,1] <- 0				# adjustment for multi-arm trials is zero for control arm
     delta[i,1] <- 0				# treatment effect is zero for control arm
     mu[i] ~ dnorm(0,.0001)		# vague priors for all trial baselines
     for (k in 1:na[i])  {			#  LOOP THROUGH ARMS
          r[i,k] ~ dbin(p[i,k],n[i,k])		# binomial likelihood
          logit(p[i,k]) <- mu[i] + delta[i,k]	# model for linear predictor
          rhat[i,k] <- p[i,k] * n[i,k]				# expected value of the numerators 
          dev[i,k] <- 2 * (r[i,k] * (log(r[i,k])-log(rhat[i,k]))     	# Deviance contribution
             +  (n[i,k]-r[i,k]) * (log(n[i,k]-r[i,k]) - log(n[i,k]-rhat[i,k])))   
        }
     resdev[i] <- sum(dev[i,1:na[i]])			# summed residual deviance contribution for this trial

     for (k in 2:na[i]) {			# LOOP THROUGH ARMS
          delta[i,k] ~ dnorm(md[i,k],taud[i,k])	    # trial-specific LOR distributions
          md[i,k] <-  d[t[i,k]] - d[t[i,1]] + sw[i,k]  # mean of LOR distributions (with multi-arm trial correction)
          taud[i,k] <- tau *2*(k-1)/k	# precision of LOR distributions (with multi-arm trial correction)
          w[i,k] <- (delta[i,k] - d[t[i,k]] + d[t[i,1]])  # adjustment for multi-arm RCTs
          sw[i,k] <- sum(w[i,1:(k-1)])/(k-1)	# cumulative adjustment for multi-arm trials # !!! w[i,1:k-1] => w[i,1:(k-1)]
         }
  }   
d[1] <- 0				# treatment effect is zero for reference treatment
for (k in 2:nt){  d[k] ~ dnorm(0,.0001) }	# vague priors for treatment effects
sd ~ dunif(0,5)				# vague prior for between-trial SD. 
tau <- pow(sd,-2)			# between-trial precision = (1/between-trial variance)

totresdev <- sum(resdev[])				#Total Residual Deviance 
}					# *** PROGRAM ENDS
" %>% textConnection



data_df <- "r[,1]	n[,1]	r[,2]	n[,2]	r[,3]	n[,3]	t[,1]	t[,2]	t[,3]	na[]	#	study ID
1472	20251	652	10396	723	10374	1	3	4	3	#	1
9	130	6	123	NA	NA	1	2	NA	2	#	2
5	63	2	59	NA	NA	1	2	NA	2	#	3
3	65	3	64	NA	NA	1	2	NA	2	#	4
887	10396	929	10372	NA	NA	1	2	NA	2	#	5
1455	13780	1418	13746	1448	13773	1	2	9	3	#	6
7	85	4	86	NA	NA	1	2	NA	2	#	7
12	159	7	157	NA	NA	1	2	NA	2	#	8
10	135	5	135	NA	NA	1	2	NA	2	#	9
4	107	6	109	NA	NA	1	4	NA	2	#	10
285	3004	270	3006	NA	NA	1	5	NA	2	#	11
11	149	2	152	NA	NA	1	7	NA	2	#	12
1	50	3	50	NA	NA	1	7	NA	2	#	13
8	58	5	54	NA	NA	1	7	NA	2	#	14
1	53	1	47	NA	NA	1	7	NA	2	#	15
4	45	0	42	NA	NA	1	7	NA	2	#	16
14	99	7	101	NA	NA	1	7	NA	2	#	17
9	41	3	46	NA	NA	1	7	NA	2	#	18
42	421	29	429	NA	NA	1	7	NA	2	#	19
2	44	3	46	NA	NA	2	7	NA	2	#	20
13	200	5	195	NA	NA	2	7	NA	2	#	21
2	56	2	47	NA	NA	2	7	NA	2	#	22
3	55	1	55	NA	NA	3	7	NA	2	#	23
10	94	3	95	NA	NA	3	7	NA	2	#	24
40	573	32	565	NA	NA	3	7	NA	2	#	25
2	61	3	62	NA	NA	3	7	NA	2	#	26
16	419	20	421	NA	NA	3	7	NA	2	#	27
5	69	3	71	NA	NA	3	7	NA	2	#	28
5	75	5	75	NA	NA	3	7	NA	2	#	29
59	782	52	790	NA	NA	3	7	NA	2	#	30
5	81	2	81	NA	NA	3	7	NA	2	#	31
16	226	12	225	NA	NA	3	7	NA	2	#	32
8	66	6	71	NA	NA	3	7	NA	2	#	33
522	8488	523	8461	NA	NA	3	6	NA	2	#	34
356	4921	757	10138	NA	NA	3	5	NA	2	#	35
13	155	7	169	NA	NA	3	5	NA	2	#	36
10	203	7	198	NA	NA	1	8	NA	2	#	37
3	58	2	52	NA	NA	1	9	NA	2	#	38
3	86	6	89	NA	NA	1	9	NA	2	#	39
3	58	2	58	NA	NA	1	9	NA	2	#	40
13	182	11	188	NA	NA	1	9	NA	2	#	41
2	26	7	54	NA	NA	3	8	NA	2	#	42
12	268	16	350	NA	NA	3	8	NA	2	#	43
5	210	17	211	NA	NA	3	9	NA	2	#	44
3	138	13	147	NA	NA	3	9	NA	2	#	45
8	132	4	66	NA	NA	2	8	NA	2	#	46
10	164	6	166	NA	NA	2	8	NA	2	#	47
6	124	5	121	NA	NA	2	8	NA	2	#	48
13	164	10	161	NA	NA	2	9	NA	2	#	49
7	93	5	90	NA	NA	2	9	NA	2	#	50" %>% 
  sub("study ID", "study_ID", .) %>% # !!! convert 'study ID' to a column instead of comments
  gsub("\\t#\\t", "\t", .) %>%
  load_dataframe_from_string

my_data <- data_df %>%
  load_matrix_data_from_df %>%
  append(list(
    ns = nrow(data_df), # number of studies
    nt = 9  # number of treatments
  ))


initial_values <- list(
  #chain 1
    list(d=c( NA, 0,0,0,0,      0,0,0,0), mu=c(0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0)),
    #chain 2
    list(d=c( NA, -1,-1,-1,-1,    -1,-1,-1,-1), mu=c(-3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,    -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3)),
    #chain 3
    list(d=c( NA, 2,5,-3,1,     -7,4,-3,2), mu=c(-3, 5,-1,-3,7,      -3, -4, -3, -3, 9,    -3, -3,-4, 3, 5,      -3, -2, 1, -3, -7,      -3,5,-1,-3,7,     -3,-4,-3,-3,0,     -3, 5,-1,-3,7,      -3, 5,-1,-3,7,      -3, -4, -3, -3, 9,    -3, -3,-4, 3, 5))
)

jags_results <- jags(
  data = my_data,
  inits = initial_values,
  parameters.to.save = c("d", "dev"),
  model.file = model_code,
  n.chains = length(initial_values),
  n.adapt = 100,
  n.iter = 60000,
  n.burnin = 10000
)

jags_summary <- summary(jags_results)


jags_summary[,c('mean', 'sd', '2.5%', '50%', '97.5%')] %>% 
  knitr::kable(., format='html', escape=FALSE, caption="JAGS Results")
```

## To Do

* Compare to WinBUGS. This will be easier if we re-sort the rows of the JAGS summary data.

# `Ch3_Crossval`

Chemo Adverse events example: 4 treatments
Predictive Cross-validation of trial 25

* Binomial likelihood, logit link, network meta-analysis (multi-arm trials)
* Random effects model with Predictive Cross-validation

This code is part of Dias, Ades, Welton, Jansen and Sutton (2018) Network Meta-Analysis for Decision Making. 
This work should be cited whenever the code is used whether in its standard form or adapted.

```{r Ch3_Crossval}
#!!! I replaced double quotes with single quotes in the code
model_code <- "model{                               # *** PROGRAM STARTS
for(i in 1:ns){                      # LOOP THROUGH STUDIES
    w[i,1] <- 0    # adjustment for multi-arm trials is zero for control arm
    delta[i,1] <- 0             # treatment effect is zero for control arm
    mu[i] ~ dnorm(0,.0001)           # vague priors for all trial baselines
    for (k in 1:na[i]) {             # LOOP THROUGH ARMS
        r[i,k] ~ dbin(p[i,k],n[i,k]) # binomial likelihood
        logit(p[i,k]) <- mu[i] + delta[i,k]  # model for linear predictor
        rhat[i,k] <- p[i,k] * n[i,k] # expected value of the numerators 
#Deviance contribution
        dev[i,k] <- 2 * (r[i,k] * (log(r[i,k])-log(rhat[i,k]))  
            +  (n[i,k]-r[i,k]) * (log(n[i,k]-r[i,k]) - log(n[i,k]-rhat[i,k])))         }
#  summed residual deviance contribution for this trial
    resdev[i] <- sum(dev[i,1:na[i]])       
    for (k in 2:na[i]) {             # LOOP THROUGH ARMS
# trial-specific LOR distributions
        delta[i,k] ~ dnorm(md[i,k],taud[i,k])
# mean of LOR distributions (with multi-arm trial correction)
        md[i,k] <-  d[t[i,k]] - d[t[i,1]] + sw[i,k]
# precision of LOR distributions (with multi-arm trial correction)
        taud[i,k] <- tau *2*(k-1)/k
# adjustment for multi-arm RCTs
        w[i,k] <- (delta[i,k] - d[t[i,k]] + d[t[i,1]])
# cumulative adjustment for multi-arm trials
        sw[i,k] <- sum(w[i,1:(k-1)])/(k-1) # !!! w[i,1:k-1] => w[i,1:(k-1)]
      }
  }   
totresdev <- sum(resdev[])           # Total Residual Deviance
d[1]<-0       # treatment effect is zero for reference treatment
# vague priors for treatment effects
for (k in 2:nt){  d[k] ~ dnorm(0,.0001) }
sd ~ dunif(0,5)     # vague prior for between-trial SD
tau <- pow(sd,-2)   # between-trial precision = (1/between-trial variance)
# predictive distribution for future trial is multivariate normal
delta.new[1] <- 0  # treatment effect is zero for reference treatment
w.new[1] <- 0      # adjustment for conditional mean is zero for ref. treat.
for (k in 2:nt) {          # LOOP THROUGH TREATMENTS
# conditional distribution of each delta.new
    delta.new[k] ~ dnorm(m.new[k],tau.new[k])
    m.new[k] <-  d[k] + sw.new[k]   # conditional mean of delta.new
    tau.new[k] <- tau *2*(k-1)/k    # conditional precision of delta.new 
    w.new[k] <- delta.new[k] - d[k] # adjustment for conditional mean
    sw.new[k] <- sum(w.new[1:(k-1)])/(k-1) # cumulative adjustment for cond. mean !!! w.new[1:k-1] => w.new[1:(k-1)]
  }
p.base ~ dbeta(a,b)       # draw baseline (control group) effect
a <- r[ns+1, 1]            # no. of events in control group
b <- n[ns+1, 1] - r[ns+1, 1]  # no. of non-events in control group
for (k in 2:na[ns+1] ) {       # LOOP THROUGH ARMS !!! 2:na[ns+1]
# predictive prob of event for each treatment arm of the new trial
    logit(p.new[k]) <- logit(p.base) 
                + (delta.new[t[ns+1,k]]- delta.new[t[ns+1,1]])
# draw predicted number of events for each arm of the new trial
    r.new[k] ~ dbin(p.new[k], n[ns+1,k])
# Bayesian p-value: probability of obtaining a value as extreme as the 
# value observed (r[ns+1,2]), given the model and the remaining data
# extreme value 'larger
#    p.cross[k] <- step(r.new[k] - r[ns+1,2]) - 0.5*equals(r.new[k],r[ns+1,2])
# extreme value 'smaller'
    p.cross[k] <- step(r[ns+1,2] - r.new[k]) - 0.5*equals(r.new[k],r[ns+1,2])
  }
# pairwise ORs and LORs for all possible pair-wise comparisons, if nt>2
for (c in 1:(nt-1)) {  
    for (k in (c+1):nt)  { 
      or[c,k] <- exp(d[k] - d[c])
      lor[c,k] <- (d[k]-d[c])
      lor.new[c,k] <- delta.new[k]- delta.new[c]
      or.new[c,k] <- exp(lor.new[c,k])
      }  
 }
}                    # *** PROGRAM ENDS   " %>% textConnection

data_df <- "t[,1]	t[,2]	na[]	r[,1]	r[,2]	n[,1]	n[,2]	#	ID
2	3	2	15	10	75	77	#	1
2	3	2	27	14	147	149	#	2
2	3	2	2	5	25	46	#	3
2	3	2	6	6	31	29	#	4
2	3	2	1	0	13	14	#	5
1	2	2	26	34	72	276	#	6
1	2	2	17	9	39	41	#	7
1	2	2	15	4	72	77	#	8
1	2	2	86	72	192	197	#	9
1	2	2	52	34	104	101	#	10
1	2	2	62	40	125	125	#	11
1	2	2	27	16	85	90	#	12
1	2	2	80	38	104	95	#	13
1	2	2	34	17	64	65	#	14
1	2	2	38	25	130	129	#	15
1	4	2	18	5	28	23	#	16
1	4	2	42	36	59	61	#	17
1	4	2	15	5	26	22	#	18
1	4	2	62	52	80	82	#	19
1	4	2	14	5	43	43	#	20
1	3	2	27	11	73	73	#	21
1	3	2	34	14	343	343	#	22
1	3	2	5	4	29	30	#	23
1	3	2	10	3	118	123	#	24
1	3	2	78	6	465	463	#	25" %>% 
  gsub("\\t#\\t", "\t", .) %>%# !!! convert 'ID' to a column instead of comments
  load_dataframe_from_string

my_data <- data_df %>%
  load_matrix_data_from_df %>%
  append(list(
    ns = nrow(data_df) - 1, # number of studies !!! MINUS ONE !!!
    nt = 4              # number of treatments
  ))


# Initial Values 

initial_values <- list(
  #chain 1
  list(d=c( NA, 0,0,0), sd=1, 
  mu=c(0,0,0,0,0,    0,0,0,0,0,    0,0,0,0,0,     0,0,0,0,0,     0,0,0,0), p.base=0.5),
  #chain 2
  list(d=c( NA, -1,-2,1), sd=4, 
  mu=c(-3,-3,-3,-3,-3,    -3,-3,-3,-3,-3,    -3,-3,-3,-3,-3,    -3,-3,-3,-3,-3,    -3,-3,-3,-3), p.base=.2),
  #chain 3
  list(d=c( NA, 2,3,-3), sd=2, 
  mu=c(-3,5,-1,-3,7,    -3,-4,-3,-3,0,    -3,5,-1,-3,7,    -3,-4,-3,-3,0,    -3,-3,0,3), p.base=.8)
)


jags_results <- jags(
  data = my_data,
  inits = initial_values,
  parameters.to.save = c("d", "delta", "dev", "lor", "or", "sd", "totresdev"),
  model.file = model_code,
  n.chains = length(initial_values),
  n.adapt = 100,
  n.iter = 60000,
  n.burnin = 10000
)

jags_summary <- summary(jags_results)


jags_summary[,c('mean', 'sd', '2.5%', '50%', '97.5%')] %>% 
  knitr::kable(., format='html', escape=FALSE, caption="JAGS Results")

```


```{r Ch3_Crossval_compare_to_WinBUGS, echo=FALSE}
"
 Results
Convergence at 60,000; 3 chains 

	 node	 mean	 sd	 MC error	2.5%	median	97.5%	start	sample
	d[2]	-0.8753	0.1389	7.271E-4	-1.161	-0.871	-0.6126	60001	300000
	d[3]	-1.178	0.2188	0.001277	-1.613	-1.177	-0.7503	60001	300000
	d[4]	-1.042	0.2653	0.001785	-1.581	-1.038	-0.5324	60001	300000
	delta[1,2]	-0.3724	0.2865	0.001281	-0.955	-0.3666	0.1869	60001	300000
	delta[2,2]	-0.4975	0.2638	0.001409	-1.049	-0.4855	-0.01179	60001	300000
	delta[3,2]	-0.2259	0.3536	0.001712	-0.8849	-0.2456	0.5346	60001	300000
	delta[4,2]	-0.2365	0.332	0.001584	-0.8666	-0.2505	0.4632	60001	300000
	delta[5,2]	-0.3496	0.3808	0.001513	-1.139	-0.3435	0.41	60001	300000
	delta[6,2]	-1.102	0.2489	0.001774	-1.627	-1.086	-0.6719	60001	300000
	delta[7,2]	-0.9125	0.2742	9.985E-4	-1.492	-0.8974	-0.3859	60001	300000
	delta[8,2]	-1.033	0.3018	0.001574	-1.705	-1.0	-0.512	60001	300000
	delta[9,2]	-0.5498	0.1982	0.001468	-0.9139	-0.5553	-0.1498	60001	300000
	delta[10,2]	-0.7807	0.2111	7.493E-4	-1.191	-0.7854	-0.3464	60001	300000
	delta[11,2]	-0.8039	0.1989	6.901E-4	-1.196	-0.8059	-0.4035	60001	300000
	delta[12,2]	-0.8339	0.2384	7.697E-4	-1.314	-0.8322	-0.3565	60001	300000
	delta[13,2]	-1.203	0.2713	0.002273	-1.772	-1.189	-0.7324	60001	300000
	delta[14,2]	-0.9814	0.252	0.001201	-1.521	-0.9623	-0.5235	60001	300000
	delta[15,2]	-0.7192	0.2201	8.927E-4	-1.134	-0.7306	-0.2577	60001	300000
	delta[16,2]	-1.205	0.365	0.002355	-1.988	-1.181	-0.5482	60001	300000
	delta[17,2]	-0.8649	0.2944	0.001712	-1.427	-0.872	-0.2697	60001	300000
	delta[18,2]	-1.138	0.3562	0.002084	-1.893	-1.119	-0.4794	60001	300000
	delta[19,2]	-0.9025	0.2773	0.001634	-1.441	-0.9058	-0.3485	60001	300000
	delta[20,2]	-1.101	0.3388	0.001929	-1.81	-1.088	-0.4636	60001	300000
	delta[21,2]	-1.191	0.2745	0.001169	-1.746	-1.187	-0.6536	60001	300000
	delta[22,2]	-1.09	0.2519	0.001172	-1.582	-1.092	-0.5856	60001	300000
	delta[23,2]	-1.058	0.3504	0.001594	-1.713	-1.077	-0.306	60001	300000
	delta[24,2]	-1.208	0.3275	0.001353	-1.882	-1.2	-0.5691	60001	300000
	delta.new[2]	-0.8755	0.3522	9.599E-4	-1.618	-0.8634	-0.1671	60001	300000
	delta.new[3]	-1.179	0.3902	0.001413	-1.972	-1.177	-0.3833	60001	300000
	delta.new[4]	-1.043	0.4186	0.001916	-1.913	-1.032	-0.2307	60001	300000
	dev[1,1]	0.8284	1.162	0.002638	8.22E-4	0.3803	4.139	60001	300000
	dev[1,2]	0.7074	1.018	0.002235	6.857E-4	0.3176	3.582	60001	300000
	dev[2,1]	1.153	1.554	0.004264	0.00121	0.5546	5.573	60001	300000
	dev[2,2]	1.065	1.418	0.004048	0.001167	0.5225	5.072	60001	300000
	dev[3,1]	0.6758	0.9396	0.002381	7.041E-4	0.3156	3.337	60001	300000
	dev[3,2]	0.9753	1.312	0.002521	0.001047	0.4682	4.702	60001	300000
	dev[4,1]	0.7297	1.034	0.002181	6.891E-4	0.3305	3.669	60001	300000
	dev[4,2]	0.7862	1.059	0.002231	8.53E-4	0.3768	3.793	60001	300000
	dev[5,1]	1.447	1.838	0.003378	0.001741	0.7641	6.545	60001	300000
	dev[5,2]	0.9088	0.9474	0.001796	0.02214	0.6113	3.483	60001	300000
	dev[6,1]	1.417	1.685	0.00875	0.001975	0.8056	6.02	60001	300000
	dev[6,2]	1.136	1.555	0.004223	0.0012	0.5376	5.565	60001	300000
	dev[7,1]	0.7327	1.041	0.002063	7.01E-4	0.3319	3.684	60001	300000
	dev[7,2]	0.6197	0.8844	0.001698	5.922E-4	0.2799	3.137	60001	300000
	dev[8,1]	1.15	1.534	0.003266	0.001279	0.5607	5.498	60001	300000
	dev[8,2]	1.072	1.273	0.003497	0.001656	0.6211	4.544	60001	300000
	dev[9,1]	1.421	1.851	0.008927	0.001541	0.7079	6.632	60001	300000
	dev[9,2]	1.477	1.893	0.008935	0.001689	0.7542	6.799	60001	300000
	dev[10,1]	0.8403	1.189	0.002621	8.222E-4	0.3808	4.217	60001	300000
	dev[10,2]	0.8213	1.158	0.002512	8.02E-4	0.3765	4.102	60001	300000
	dev[11,1]	0.8277	1.173	0.002622	8.253E-4	0.3758	4.172	60001	300000
	dev[11,2]	0.8098	1.153	0.002528	7.963E-4	0.366	4.092	60001	300000
	dev[12,1]	0.7751	1.101	0.002228	7.43E-4	0.3501	3.922	60001	300000
	dev[12,2]	0.695	0.9901	0.002281	6.892E-4	0.3147	3.508	60001	300000
	dev[13,1]	1.8	2.111	0.0128	0.002651	1.043	7.56	60001	300000
	dev[13,2]	1.732	2.083	0.01079	0.002348	0.9718	7.465	60001	300000
	dev[14,1]	0.8753	1.218	0.002763	9.184E-4	0.4029	4.335	60001	300000
	dev[14,2]	0.8189	1.141	0.002267	8.542E-4	0.38	4.048	60001	300000
	dev[15,1]	0.9315	1.302	0.002736	9.191E-4	0.428	4.627	60001	300000
	dev[15,2]	0.9716	1.302	0.002778	0.001059	0.4728	4.642	60001	300000
	dev[16,1]	1.219	1.581	0.004843	0.001405	0.6175	5.659	60001	300000
	dev[16,2]	1.285	1.54	0.005669	0.001854	0.735	5.503	60001	300000
	dev[17,1]	1.174	1.524	0.005593	0.001324	0.5927	5.452	60001	300000
	dev[17,2]	1.082	1.473	0.003603	0.001122	0.5136	5.259	60001	300000
	dev[18,1]	0.912	1.269	0.003177	9.329E-4	0.4225	4.514	60001	300000
	dev[18,2]	0.8144	1.122	0.003323	8.064E-4	0.3827	4.014	60001	300000
	dev[19,1]	1.014	1.376	0.004894	0.001078	0.4814	4.922	60001	300000
	dev[19,2]	0.9597	1.343	0.003048	9.723E-4	0.443	4.78	60001	300000
	dev[20,1]	0.8604	1.211	0.002796	8.543E-4	0.3952	4.304	60001	300000
	dev[20,2]	0.6337	0.9134	0.002675	6.188E-4	0.2837	3.225	60001	300000
	dev[21,1]	0.8118	1.152	0.002326	8.171E-4	0.3671	4.095	60001	300000
	dev[21,2]	0.6464	0.9296	0.002113	6.257E-4	0.29	3.281	60001	300000
	dev[22,1]	0.8967	1.268	0.002612	9.027E-4	0.4098	4.496	60001	300000
	dev[22,2]	0.8458	1.151	0.003363	8.663E-4	0.4024	4.108	60001	300000
	dev[23,1]	0.9911	1.348	0.002892	0.001027	0.4727	4.793	60001	300000
	dev[23,2]	1.262	1.382	0.003422	0.002284	0.8035	4.967	60001	300000
	dev[24,1]	0.8448	1.189	0.002264	8.26E-4	0.3854	4.226	60001	300000
	dev[24,2]	0.4403	0.6507	0.001547	4.156E-4	0.1933	2.272	60001	300000
	lor[1,2]	-0.8753	0.1389	7.271E-4	-1.161	-0.871	-0.6126	60001	300000
	lor[1,3]	-1.178	0.2188	0.001277	-1.613	-1.177	-0.7503	60001	300000
	lor[1,4]	-1.042	0.2653	0.001785	-1.581	-1.038	-0.5324	60001	300000
	lor[2,3]	-0.3029	0.2204	0.001345	-0.7318	-0.3038	0.1386	60001	300000
	lor[2,4]	-0.1671	0.2964	0.001808	-0.7612	-0.1633	0.408	60001	300000
	lor[3,4]	0.1358	0.344	0.00219	-0.5613	0.1401	0.7985	60001	300000
	lor.new[1,2]	-0.8755	0.3522	9.599E-4	-1.618	-0.8634	-0.1671	60001	300000
	lor.new[1,3]	-1.179	0.3902	0.001413	-1.972	-1.177	-0.3833	60001	300000
	lor.new[1,4]	-1.043	0.4186	0.001916	-1.913	-1.032	-0.2307	60001	300000
	lor.new[2,3]	-0.3031	0.3912	0.001488	-1.082	-0.3096	0.5104	60001	300000
	lor.new[2,4]	-0.1673	0.4379	0.0019	-1.059	-0.1626	0.7	60001	300000
	lor.new[3,4]	0.1358	0.4724	0.00227	-0.8357	0.1437	1.054	60001	300000
	or[1,2]	0.4207	0.05817	3.018E-4	0.3132	0.4185	0.542	60001	300000
	or[1,3]	0.3153	0.06984	4.026E-4	0.1992	0.3082	0.4722	60001	300000
	or[1,4]	0.3651	0.09747	6.933E-4	0.2058	0.3542	0.5872	60001	300000
	or[2,3]	0.7569	0.1707	0.00102	0.481	0.738	1.149	60001	300000
	or[2,4]	0.8839	0.2676	0.001636	0.4671	0.8494	1.504	60001	300000
	or[3,4]	1.215	0.4267	0.002675	0.5705	1.15	2.222	60001	300000
	or.new[1,2]	0.4436	0.1851	4.433E-4	0.1983	0.4217	0.8462	60001	300000
	or.new[1,3]	0.3325	0.1448	5.062E-4	0.1392	0.3081	0.6816	60001	300000
	or.new[1,4]	0.3845	0.1729	7.446E-4	0.1477	0.3564	0.794	60001	300000
	or.new[2,3]	0.7991	0.3573	0.00133	0.3388	0.7338	1.666	60001	300000
	or.new[2,4]	0.932	0.4585	0.001835	0.3467	0.8499	2.014	60001	300000
	or.new[3,4]	1.28	0.6626	0.002863	0.4336	1.155	2.868	60001	300000
	p.cross[2]	0.003887	0.05773	1.101E-4	0.0	0.0	0.0	60001	300000
	sd	0.2946	0.1324	0.001485	0.04521	0.2891	0.5758	60001	300000
	totresdev	46.89	9.219	0.05388	30.51	46.36	66.38	60001	300000
	 
# 48 unconstrained data points

Dbar = post.mean of -2logL; Dhat = -2LogL at post.mean of stochastic nodes
	Dbar	Dhat	pD	DIC	
r	245.668	211.096	34.572	280.241	
total	245.668	211.096	34.572	280.241	
"

```

