---
title: "NMADM Chapter02"
author: "Bob"
date: "2025-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE, warning=FALSE}
# libraries
library('rjags')      # coda.samples
library('jagsUI')     # wrapper for rjags
library('mcmcplots')  # caterplot
library('coda')
library('knitr')
library('dplyr')
library('tidyr')
library('ggplot2')
library('knitr')      # HTML table display
```

## Notes:

* Results the multi-comparison models have not been checked against WinBUGS yet.
* In the last example, I found a new WinBUGS :: JAGS difference I had not seen before: 
`w[i,1:k-1]` -> `w[i,1:(k-1)]`

# Ch2_FE_Bi_logit_pair

Thrombolytics example - two treatments only
treatment 1 = a-tPA
treatment 2 = PTCA

Binomial likelihood, logit link, pairwise meta-analysis (2 treatments)
Fixed effect model

This code is part of Dias, Ades, Welton, Jansen and Sutton (2018) Network Meta-Analysis for Decision Making. 
This work should be cited whenever the code is used whether in its standard form or adapted.

```{r Ch2_FE_Bi_logit_pair}
model_code <- "
model{                                 # *** PROGRAM STARTS
for(i in 1:ns){                        # LOOP THROUGH STUDIES
    mu[i] ~ dnorm(0,.0001)             # vague priors for all trial baselines
    for (k in 1:2) {                   # LOOP THROUGH ARMS
        r[i,k] ~ dbin(p[i,k],n[i,k])   # binomial likelihood
        logit(p[i,k]) <- mu[i] + d[k]  # model for linear predictor
      }
  }   
d[1]<- 0              # treatment effect is zero for reference treatment
d[2] ~ dnorm(0,.0001) # vague prior for treatment effect
or <- exp(d[2])       # calculates the odds ratio
prob.harm <- step(d[2])  # calculates probability treat 2 is harmful
}                                     # *** PROGRAM ENDS                         
" |> textConnection()

# Data (Extended Thrombolytics example - 2 treatments)

df <- "r[,1]	n[,1]	r[,2]	n[,2]	Study_ID
3	55	1	55	Ribichini_1996
10	94	3	95	Garcia_1997
40	573	32	565	GUSTO-2_1997
5	75	5	75	Vermeer_1999
5	69	3	71	Schomig_2000
2	61	3	62	LeMay_2001
19	419	20	421	Bonnefoy_2002
59	782	52	790	Andersen_2002
5	81	2	81	Kastrati_2002
16	226	12	225	Aversano_2002
8	66	6	71	Grines_2002" |> textConnection() |> read.delim()


data <- list(
  r = cbind(df[,1], df[,3]),
  n = cbind(df[,2], df[,4]),
  ns = nrow(df) # ns= number of studies
)   



# Initial Values
initial_values = list(
  #chain 1
  list(d=c( NA, 0),  mu=c(0,0,0,0,0,     0,0,0,0,0,     0)),
  #chain 2
  list(d=c( NA, -1), mu=c(-3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3)),
  #chain 3
  list(d=c( NA, 2), mu=c(-3,5,-1,-3,7,     -3,-4,-3,-3,0,     -7))
)

results <- jags(
  data = data,
  inits = initial_values,
  parameters.to.save = c("d", "or", "prob.harm"), # Monitor: d, or, prob.harm
  model.file = model_code,
  n.chains = length(initial_values),
  n.adapt = 100,
  n.iter = 60000,
  n.burnin = 10000
)

results_summary <- summary(results)

results_summary[c('d[2]', 'or', 'prob.harm'), c("mean", "sd", "2.5%", "50%", "97.5%")] %>% 
  knitr::kable(., format='html', escape=FALSE, caption="JAGS Results")

#  Results
# Convergence at 10,000; 3 chains 




WinBUGS_results_df <- "node	 mean	 sd	 MC_error	2.5%	median	97.5%	start	sample
d[2]	-0.2336	0.1178	7.668E-4	-0.4654	-0.2333	-0.002876	10001	60000
or	0.7971	0.09416	6.144E-4	0.6279	0.7919	0.9971	10001	60000
prob.harm	0.02362	0.1519	7.52E-4	0.0	0.0	0.0	10001	60000" %>% 
  gsub("[ |\t]+", " ", .) %>% # !!! tabs get converted to one or more spaces in RStudio; make them all one space
  textConnection() %>% 
  read.delim(sep=' ', row.names=1)

WinBUGS_results_df %>% 
  knitr::kable(., format='html', escape=FALSE, caption="WinBUGS Results")

```

# Ch2_FE_Bi_logit

Extended Thrombolytics example (Caldwell et al, BMJ 2005)

```
=============================
Treatment codes
1	SK
2	t-PA
3	at-PA
4	SK+t-PA
5	r-PA
6	TNK
7	PTCA
=============================
```

Binomial likelihood, logit link
Fixed effect model 

This code is part of Dias, Ades, Welton, Jansen and Sutton (2018) Network Meta-Analysis for Decision Making. 
This work should be cited whenever the code is used whether in its standard form or adapted.


```{r Ch2_FE_Bi_logit}
model_code <- "
data{
  smidge <- 2.225074e-308 # !!! kludge alert !!! add a smidge to denominator to avoid dividing by zero.
}

model{
  for(i in 1:ns){                 # LOOP THROUGH STUDIES
    mu[i] ~ dnorm(0,.0001)        # vague priors for all trial baselines
    for (k in 1:na[i])  {         # LOOP THROUGH ARMS
        r[i,k] ~ dbin(p[i,k],n[i,k])    # binomial likelihood
        # model for linear predictor
        logit(p[i,k]) <- mu[i] + d[t[i,k]] - d[t[i,1]]
    }
  }   
  d[1] <- 0    # treatment effect is zero for reference treatment
  # vague priors for treatment effects
  for (k in 2:nt){  d[k] ~ dnorm(0,.0001) }
  # pairwise ORs and LORs for all possible pair-wise comparisons
  for (c in 1:(nt-1)) {  
    for (k in (c+1):nt)  { 
      or[c,k] <- exp(d[k] - d[c])
      lor[c,k] <- (d[k] - d[c])
    }  
  }
  # ranking on relative scale
  rk <-  rank(d[]) # !!! Bob
  for (k in 1:nt) { 
      # rk[k] <- nt+1-rank(d[],k)	# assumes events are 'good'
      # rk[k] <- rank(d[],k)	# assumes events are 'bad' # !!! rank function different in JAGS
      best[k] <- equals(rk[k],1)	#calculate probability that treat k is best
      # calculates probability that treat k is h-th best
      for (h in 1:nt){  prob[h,k] <- equals(rk[k],h)  }
   }
  # Provide estimates of treatment effects T[k] on the natural (probability) scale 
  # Given a Mean Effect, meanA, for treatment 1, with precision (1/variance) precA
  A ~ dnorm(meanA,precA)
  for (k in 1:nt) { logit(T[k]) <- A + d[k]  }
  # Provide estimates of number needed to treat NNT[k], Risk Difference RD[k], 
  # and Relative Risk RR[k], for each treatment, relative to treatment 1
  for (k in 2:nt) { 
    # NNT[k] <- 1/(T[k] - T[1] + smidge)	# assumes events are 'good'
    NNT[k] <- 1/(T[1] - T[k] + smidge)	  # assumes events are 'bad'
    RD[k] <- T[k] - T[1]
    RR[k] <- T[k]/T[1]
  }
}" |> textConnection()


load_data_from_string <- function(data_str){
  # R does not like solumn names with special characters like square brackets,
  # but we can re-set them to their original values.
  column_names <- data_str %>% strsplit("\\n") %>% "[["(1) %>% "[["(1) %>% strsplit("\\t") %>% "[["(1)
  df <- data_str |> textConnection() |> read.delim(row.names=NULL) %>% setNames(nm=column_names)
}


df <- "na[]	t[,1]	t[,2]	t[,3]	r[,1]	n[,1]	r[,2]	n[,2]	r[,3]	n[,3]	ID	year
3	1	3	4	1472	20251	652	10396	723	10374	GUSTO-1	1993
2	1	2	NA	3	65	3	64	NA	NA	ECSG	1985
2	1	2	NA	12	159	7	157	NA	NA	TIMI-1	1987
2	1	2	NA	7	85	4	86	NA	NA	PAIMS	1989
2	1	2	NA	10	135	5	135	NA	NA	White	1989
2	1	2	NA	887	10396	929	10372	NA	NA	GISSI-2	1990
2	1	2	NA	5	63	2	59	NA	NA	Cherng	1992
2	1	2	NA	1455	13780	1418	13746	NA	NA	ISIS-3	1992
2	1	2	NA	9	130	6	123	NA	NA	CI	1993
2	1	4	NA	4	107	6	109	NA	NA	KAMIT	1991
2	1	5	NA	285	3004	270	3006	NA	NA	INJECT	1995
2	1	7	NA	11	149	2	152	NA	NA	Zijlstra	1993
2	1	7	NA	1	50	3	50	NA	NA	Riberio	1993
2	1	7	NA	8	58	5	54	NA	NA	Grinfeld	1996
2	1	7	NA	1	53	1	47	NA	NA	Zijlstra	1997
2	1	7	NA	4	45	0	42	NA	NA	Akhras	1997
2	1	7	NA	14	99	7	101	NA	NA	Widimsky	2000
2	1	7	NA	9	41	3	46	NA	NA	DeBoer	2002
2	1	7	NA	42	421	29	429	NA	NA	Widimsky	2002
2	2	7	NA	2	44	3	46	NA	NA	DeWood	1990
2	2	7	NA	13	200	5	195	NA	NA	Grines	1993
2	2	7	NA	2	56	2	47	NA	NA	Gibbons	1993
2	3	5	NA	13	155	7	169	NA	NA	RAPID-2	1996
2	3	5	NA	356	4921	757	10138	NA	NA	GUSTO-3	1997
2	3	6	NA	522	8488	523	8461	NA	NA	ASSENT-2	1999
2	3	7	NA	3	55	1	55	NA	NA	Ribichini	1996
2	3	7	NA	10	94	3	95	NA	NA	Garcia	1997
2	3	7	NA	40	573	32	565	NA	NA	GUSTO-2	1997
2	3	7	NA	5	75	5	75	NA	NA	Vermeer	1999
2	3	7	NA	5	69	3	71	NA	NA	Schomig	2000
2	3	7	NA	2	61	3	62	NA	NA	LeMay	2001
2	3	7	NA	19	419	20	421	NA	NA	Bonnefoy	2002
2	3	7	NA	59	782	52	790	NA	NA	Andersen	2002
2	3	7	NA	5	81	2	81	NA	NA	Kastrati	2002
2	3	7	NA	16	226	12	225	NA	NA	Aversano	2002
2	3	7	NA	8	66	6	71	NA	NA	Grines	2002" %>% load_data_from_string


data <- list(
  na = df$na[],
  t = cbind(df[['t[,1]']], df[['t[,2]']], df[['t[,3]']]),
  r = cbind(df[['r[,1]']], df[['r[,2]']], df[['r[,3]']]),
  n = cbind(df[['n[,1]']], df[['n[,2]']], df[['n[,3]']]),
  ns=36, # number of studies
  nt=7,  # number of treatments
  meanA= -2.39, 
  precA=11.9)

# Initial Values 
initial_values <- list(
  #chain 1
  list(d=c( NA, 0,0,0,0,      0,0), mu=c(0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0)),
  #chain 2
  list(d=c( NA, -1,-1,-1,-1,    -1,-1), mu=c(-3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3)),
  #chain 3
  list(d=c( NA, 2,5,-3,1,     -7,4), mu=c(-3, 5,-1,-3,7,      -3, -4, -3, -3, 9,    -3, -3,-4, 3, 5,      -3, -2, 1, -3, -7,      -3,5,-1,-3,7,     -3,-4,-3,-3,0,     -3, 5,-1,-3,7,      -7))
)

results <- jags(
  data = data,
  inits = initial_values,
  parameters.to.save = c("d", "or", "NNT", "RD", "RR"),
  model.file = model_code,
  n.chains = length(initial_values),
  n.adapt = 100,
  n.iter = 60000,
  n.burnin = 10000
)

summary(results)

```

I have not checked these results against WinBUGS yet.


# Ch2_RE_Bi_logit_pair

Thrombolytics example - two treatments only
treatment 1 = a-tPA
treatment 2 = PTCA

Binomial likelihood, logit link, pairwise meta-analysis (2 treatments)
Random effects model

This code is part of Dias, Ades, Welton, Jansen and Sutton (2018) Network Meta-Analysis for Decision Making. 
This work should be cited whenever the code is used whether in its standard form or adapted.


```{r Ch2_RE_Bi_logit_pair}

model_code <- "model{             # *** PROGRAM STARTS
  for(i in 1:ns){                 # LOOP THROUGH STUDIES
    delta[i,1] <- 0               # treatment effect is zero for control arm
    mu[i] ~ dnorm(0,.0001)        # vague priors for all trial baselines
    for (k in 1:2) {              # LOOP THROUGH ARMS
      r[i,k] ~ dbin(p[i,k],n[i,k])        # binomial likelihood
      logit(p[i,k]) <- mu[i] + delta[i,k] # model for linear predictor
    }
    delta[i,2] ~ dnorm(d[2],tau)  # trial-specific LOR distributions
  }
  d[1]<- 0                        # treatment effect is zero for reference treatment
  d[2] ~ dnorm(0,.0001)           # vague prior for treatment effect
  sd ~ dunif(0,2)                 # vague prior for between-trial SD
  tau <- pow(sd,-2)               # between-trial precision = (1/between-trial variance)
  OR <- exp(d[2])                 # calculates the odds ratio
  prob.harm <- step(d[2])         # calculates probability treat 2 is harmful
}                                 # *** PROGRAM ENDS
" |> textConnection()

df <- "r[,1]	n[,1]	r[,2]	n[,2]\tID\tyear
3	55	1	55	Ribichini	1996
10	94	3	95	Garcia	1997
40	573	32	565	GUSTO-2	1997
5	75	5	75	Vermeer	1999
5	69	3	71	Schomig	2000
2	61	3	62	LeMay	2001
19	419	20	421	Bonnefoy	2002
59	782	52	790	Andersen	2002
5	81	2	81	Kastrati	2002
16	226	12	225	Aversano	2002
8	66	6	71	Grines	2002" %>% load_data_from_string

data <- list(
  r = cbind(df[['r[,1]']], df[['r[,2]']]),
  n = cbind(df[['n[,1]']], df[['n[,2]']]),
  ns = nrow(df) # number of studies
) 

initial_values <- list(
  # chain 1
  list(d=c( NA, 0), sd=1, mu=c(0,0,0,0,0,     0,0,0,0,0,     0)),
  # chain 2
  list(d=c( NA, -1), sd=0.1, mu=c(-3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3)),
  # chain 3
  list(d=c( NA, 2), sd=0.5, mu=c(-3,5,-1,-3,7,     -3,-4,-3,-3,0,     -7))
)

results <- jags(
  data = data,
  # inits = initial_values,
  parameters.to.save = c("d", "sd", "OR", "prob.harm"),
  model.file = model_code,
  n.chains = length(initial_values),
  n.adapt = 100,
  n.iter = 60000,
  n.burnin = 10000
)

summary(results)[,c('mean', 'sd', '2.5%', '50%', '97.5%')] %>% 
  knitr::kable(., format='html', escape=FALSE, caption="JAGS Results")


WinBUGS_results_df <- "node	 mean	 sd	 MC error	2.5%	median	97.5%	start	sample
OR	0.7724	0.1197	0.002062	0.5485	0.7684	1.022	10001	60000
d[2]	-0.2705	0.1576	0.002687	-0.6006	-0.2635	0.02178	10001	60000
prob.harm	0.03452	0.1826	0.002824	0.0	0.0	1.0	10001	60000
sd	0.1966	0.1734	0.003771	0.007339	0.1513	0.6419	10001	60000" %>%
  load_data_from_string 

WinBUGS_results_df %>%
  knitr::kable(., format='html', escape=FALSE, caption="WinBUGS Results")
```


# Ch2_RE_Bi_logit

Extended Thrombolytics example (Caldwell et al, BMJ 2005)

```
=============================
Treatment codes
1	SK
2	t-PA
3	at-PA
4	SK+t-PA
5	r-PA
6	TNK
7	PTCA
============================= 
```

Binomial likelihood, logit link
Random effects model for multi-arm trials

This code is part of Dias, Ades, Welton, Jansen and Sutton (2018) Network Meta-Analysis for Decision Making. 
This work should be cited whenever the code is used whether in its standard form or adapted.


```{r Ch2_RE_Bi_logit}

model_code <- "
data{
  smidge <- 2.225074e-308 # !!! kludge alert !!! add a smidge to denominator to avoid dividing by zero.
}

model{                                 # *** PROGRAM STARTS
  for(i in 1:ns){                      # LOOP THROUGH STUDIES
      w[i,1] <- 0                      # adjustment for multi-arm trials is zero for control arm
      delta[i,1] <- 0                  # treatment effect is zero for control arm
      mu[i] ~ dnorm(0,.0001)           # vague priors for all trial baselines
      for (k in 1:na[i]) {             # LOOP THROUGH ARMS
          r[i,k] ~ dbin(p[i,k],n[i,k]) # binomial likelihood
          logit(p[i,k]) <- mu[i] + delta[i,k]  # model for linear predictor
      }
      for (k in 2:na[i]) {             # LOOP THROUGH ARMS
          # trial-specific LOR distributions
          delta[i,k] ~ dnorm(md[i,k], taud[i,k])
          # mean of LOR distributions (with multi-arm trial correction)
          md[i,k] <-  d[t[i,k]] - d[t[i,1]] + sw[i,k]
          # precision of LOR distributions (with multi-arm trial correction)
          taud[i,k] <- tau *2*(k-1)/k
          # adjustment for multi-arm RCTs
          w[i,k] <- (delta[i,k] - d[t[i,k]] + d[t[i,1]])
          # cumulative adjustment for multi-arm trials
          sw[i,k] <- sum(w[i,1:(k-1)])/(k-1)  # !!! Index out of range error: put parentheses around (k-1)
      }
  }   
  d[1]<-0             # treatment effect is zero for reference treatment
  # vague priors for treatment effects
  for (k in 2:nt){  d[k] ~ dnorm(0,.0001) }
  sd ~ dunif(0,2)     # vague prior for between-trial SD
  tau <- pow(sd,-2)   # between-trial precision = (1/between-trial variance)
  # pairwise ORs and LORs for all possible pair-wise comparisons
  for (c in 1:(nt-1)) {  
    for (k in (c+1):nt)  { 
      or[c,k] <- exp(d[k] - d[c])
      lor[c,k] <- (d[k]-d[c])
    }  
  }
  # ranking on relative scale
  rank_d <- rank(d)
  for (k in 1:nt) { 
    # rk[k] <- nt+1-rank_d[k]	# assumes events are 'good'; !!! was 'nt+1-rank(d[],k)' in WinBugs
    rk[k] <- rank_d[k]	# assumes events are 'bad' !!! was 'rank(d[],k)' in WinBugs
    best[k] <- equals(rk[k],1)	#calculate probability that treat k is best
    # calculates probability that treat k is h-th best
    for (h in 1:nt){  prob[h,k] <- equals(rk[k],h)}
  }
  # Provide estimates of treatment effects T[k] on the natural (probability) scale 
  # Given a Mean Effect, meanA, for treatment 1, with precision (1/variance) precA
  A ~ dnorm(meanA,precA)
  for (k in 1:nt) { logit(T[k]) <- A + d[k]  }
  # Provide estimates of number needed to treat NNT[k], Risk Difference RD[k], 
  # and Relative Risk RR[k], for each treatment, relative to treatment 1
  for (k in 2:nt) { 
    # NNT[k] <- 1/(T[k] - T[1] + smidge)	# assumes events are 'good'
    NNT[k] <- 1/(T[1] - T[k] + smidge)	# assumes events are 'bad'
    RD[k] <- T[k] - T[1]
    RR[k] <- T[k]/T[1]
  }
}                                    # *** PROGRAM ENDS                          
" %>% textConnection()

df <- "na[]	t[,1]	t[,2]	t[,3]	r[,1]	n[,1]	r[,2]	n[,2]	r[,3]	n[,3]	#	ID	year
3	1	3	4	1472	20251	652	10396	723	10374	#	GUSTO-1	1993
2	1	2	NA	3	65	3	64	NA	NA	#	ECSG	1985
2	1	2	NA	12	159	7	157	NA	NA	#	TIMI-1	1987
2	1	2	NA	7	85	4	86	NA	NA	#	PAIMS	1989
2	1	2	NA	10	135	5	135	NA	NA	#	White	1989
2	1	2	NA	887	10396	929	10372	NA	NA	#	GISSI-2	1990
2	1	2	NA	5	63	2	59	NA	NA	#	Cherng	1992
2	1	2	NA	1455	13780	1418	13746	NA	NA	#	ISIS-3	1992
2	1	2	NA	9	130	6	123	NA	NA	#	CI	1993
2	1	4	NA	4	107	6	109	NA	NA	#	KAMIT	1991
2	1	5	NA	285	3004	270	3006	NA	NA	#	INJECT	1995
2	1	7	NA	11	149	2	152	NA	NA	#	Zijlstra	1993
2	1	7	NA	1	50	3	50	NA	NA	#	Riberio	1993
2	1	7	NA	8	58	5	54	NA	NA	#	Grinfeld	1996
2	1	7	NA	1	53	1	47	NA	NA	#	Zijlstra	1997
2	1	7	NA	4	45	0	42	NA	NA	#	Akhras	1997
2	1	7	NA	14	99	7	101	NA	NA	#	Widimsky	2000
2	1	7	NA	9	41	3	46	NA	NA	#	DeBoer	2002
2	1	7	NA	42	421	29	429	NA	NA	#	Widimsky	2002
2	2	7	NA	2	44	3	46	NA	NA	#	DeWood	1990
2	2	7	NA	13	200	5	195	NA	NA	#	Grines	1993
2	2	7	NA	2	56	2	47	NA	NA	#	Gibbons	1993
2	3	5	NA	13	155	7	169	NA	NA	#	RAPID-2	1996
2	3	5	NA	356	4921	757	10138	NA	NA	#	GUSTO-3	1997
2	3	6	NA	522	8488	523	8461	NA	NA	#	ASSENT-2	1999
2	3	7	NA	3	55	1	55	NA	NA	#	Ribichini	1996
2	3	7	NA	10	94	3	95	NA	NA	#	Garcia	1997
2	3	7	NA	40	573	32	565	NA	NA	#	GUSTO-2	1997
2	3	7	NA	5	75	5	75	NA	NA	#	Vermeer	1999
2	3	7	NA	5	69	3	71	NA	NA	#	Schomig	2000
2	3	7	NA	2	61	3	62	NA	NA	#	LeMay	2001
2	3	7	NA	19	419	20	421	NA	NA	#	Bonnefoy	2002
2	3	7	NA	59	782	52	790	NA	NA	#	Andersen	2002
2	3	7	NA	5	81	2	81	NA	NA	#	Kastrati	2002
2	3	7	NA	16	226	12	225	NA	NA	#	Aversano	2002
2	3	7	NA	8	66	6	71	NA	NA	#	Grines	2002" %>% load_data_from_string

data <- list(
  na = df[['na[]']],
  t = cbind(df[['t[,1]']], df[['t[,2]']], df[['t[,3]']]),
  r = cbind(df[['r[,1]']], df[['r[,2]']], df[['r[,3]']]),
  n = cbind(df[['n[,1]']], df[['n[,2]']], df[['n[,3]']]),
  ns = nrow(df), # number of studies
  nt = 7,  # number of treatments
  meanA = -2.39, 
  precA = 11.9
)

initial_values <- list(
  # chain 1
  list(d=c( NA, 0,0,0,0,      0,0), sd=1,
mu=c(0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0,0,0,0,0,     0)),

  # chain 2
  list(d=c( NA, -1,-1,-1,-1,    -1,-1), sd=.5,
  mu=c(-3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3,-3,-3,-3,-3,     -3)),

  # chain 3
  list(d=c( NA, 2,5,-3,1,     -7,4), sd=.1,
  mu=c(-3, 5,-1,-3,7,      -3, -4, -3, -3, 9,    -3, -3,-4, 3, 5,      -3, -2, 1, -3, -7,      -3,5,-1,-3,7,     -3,-4,-3,-3,0,     -3, 5,-1,-3,7,      -7))
)


results <- jags(
  data = data,
  # inits = initial_values,
  parameters.to.save = c("d", "sd", "OR", "prob.harm"),
  model.file = model_code,
  n.chains = length(initial_values),
  n.adapt = 100,
  n.iter = 60000,
  n.burnin = 10000
)

summary(results)[,c('mean', 'sd', '2.5%', '50%', '97.5%')] %>% 
  knitr::kable(., format='html', escape=FALSE, caption="JAGS Results")

```

These results have not been checked against WinBUGS yet.
